Convex Surrogates via Polyhedral Losses

### 2019-02-26 jessie-writeup/n9-not-3-embed.tex
Adding post-submission notes in here

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{mathtools, amsmath, amsthm, amssymb, graphicx, mathrsfs, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}
\usepackage[margin=1.25in]{geometry}

@@

\newcommand{\inter}[1]{\mathring{#1}}%\mathrm{int}(#1)}
\newcommand{\cl}[1]{\text{cl}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\ones}{\mathbbm{1}}
\newcommand{\aff}{\text{aff}}
\newcommand{\im}{\text{im}}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\card}{\textbf{card}}

@@

>>>>

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{mathtools, amsmath, amsthm, amssymb, graphicx, mathrsfs, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}
\usepackage{tikzsymbols}
\usepackage[margin=1.25in]{geometry}

@@

\newcommand{\inter}[1]{\mathring{#1}}%\mathrm{int}(#1)}
\newcommand{\cl}[1]{\text{cl}(#1)}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\inprod}[2]{\langle #1, #2 \rangle}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\ones}{\mathbbm{1}}
\newcommand{\aff}{\text{aff}}
\newcommand{\im}{\text{im}}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\card}{\textbf{card}}

@@

\newpage
\section{Post-submission thoughts}
\subsection{The mode}
\jessiet{For all the $\leq$ statements in here, there might be a reason they would have to be equalities, but I haven't been able to think it fully through/prove it, so they are in here as inequalities for now.}
\begin{conjecture}
	The normals constructing a tight H-rep for a cell for the mode must be linearly independent
\end{conjecture}
\begin{proof}
  Consider the matrix describing the cell for outcome $n$...

  $B^{(n)} := \begin{pmatrix}
  -1 & 0 & 0 & \ldots & 1 \\
  0 & -1 & 0 & \ldots & 1 \\
  0 & 0 & -1 & \ldots &  1\\
  0 & 0 & \ldots &-1 &1\\
  \end{pmatrix}$

  Now for each column $B_i^{(n)}$ of $B^{(n)}$, we must have a set of normals such that for each normal $j$, we must have $\inprod{V_j}{x} = B_{i,j}$ in order for the halfspace representation to be tight.
  Let us write the polytope $T_i := \{x \in \reals^d : Vx \leq B_i \}$ as the feasible polytope for the H-rep of the property.
  In order for $\{x \in T_n : \inprod{V_j}{x} = 1 \}$ to be nonempty for all $V_j$, we assert the normals $V_j$ must all be linearly independent.

%	The first $n-1$ columns of $B^{(n)}$ form the standard basis in $\reals^{n-1}$.

  To see this, first consider that if $V_2 = -V_1$ , then one of these constraints for the polytope $T_n$ is redundant.
  Therefore these normals do not form a tight $H$-rep since for (WLOG) $V_2$, the halfspace represented by $\{x : \inprod{V_2}{x} \leq 1 \}$ is a (strict since none of the normals can be $0$) subset of the halfspace $\{x : \inprod{V_1}{x} \leq 1 \}$.

  To generalize this to linear independence of the normals, suppose $V_{\Laughey} = \sum_{i \neq \Laughey} \lambda_i V_i$ so that $V_{\Laughey}$ is a linear combination of other normals.
%  
  The tightness condition, for the columns being indexed by $i$ and rows indexed by $j$ is as follows:
  For all $i \in \{1, 2, \ldots, n\}$ and $j \in \{1, 2, \ldots, k = n-1\}$ there must be an $x \in T_i$ so that 
  \begin{align*}
 \inprod{V_j}{x} = 
  \begin{cases}
  1 & i = n\\
  -1& i = j\\
  0 & i \neq j, i \neq n
  \end{cases}
  \end{align*}

  Given the tightness constraints, if $V_{\Laughey}$ is a linear combination of $V_i$s, then $\inprod{V_{\Laughey}}{x} = \sum_{i\neq \Laughey} \inprod{\lambda_i V_i}{x} \leq 0 + 0 + \ldots + 0 + \lambda_n = \lambda_n$.

  If $\lambda_n= -1$, we contradict the tightness of $T_n$ shown above. 
  Otherwise, we contradict the tightness of $T_i$ since there is no constraint such that $\inprod{V_{\Laughey}}{x} = -1$ given the tightness of the other constraints.

\begin{corollary}
  The mode is not $(n-2)$ embeddable
\end{corollary}
\begin{proof}
  In order to have a tight $H$-rep, we need normals that are linearly independent.
  Since the normals must be linearly independent, it must be true that for $V \in \reals^{k \times d}$ that $k \leq d$.
  However, since $k = n-1 \leq d$, we know that the embedding dimension must be at least $n-1$.
\end{proof}

\subsection{Abstain}
Let's think about abstain for a second in a similar mindset.
\begin{theorem}
	The normals constructing a tight H-rep for the abstain(1/2) property on $n$ outcomes \jessie{must sum to $\vec{0}$ and} [[COOL CHARACTERIZATION]]
\end{theorem}
\begin{proof}
  Consider the matrix $B^{(abs)} \in \reals^{k \times n}$ (where $k = n$) describing the abstain cell...

  $B^{(abs)} := \begin{pmatrix}
  -1 & 1 & 1 & \ldots & 1\\
  1 & -1 & 1 & \ldots & 1\\
  1 & 1 & -1 & \ldots & 1\\
  1 & 1 & \ldots & 1 & -1\\
  \end{pmatrix}$

  The tightness constraint for the abstain cell is for all columns of $B^{(abs)}$, indexed by $i$, and rows of $V$, indexed by $j$, we know there must be an $x \in T_i$ such that 
  \begin{align*}
  \inprod{V_j}{x} = 
  \begin{cases}
  -1 & i = j\\
  1& o/w\\
  \end{cases}
  \end{align*}

  When a normal $V_{\Laughey}$ can be written as a linear combination of other normals $\sum_{j \neq \Laughey} \lambda_j V_j$, then we know $\inprod{V_{\Laughey}}{x} = \sum_{j \neq \Laughey} \inprod{V_j}{x}$.
  Further, by the requirement that $x \in T_i$, we know that
  \begin{align*}
  	\inprod{V_{\Laughey}}{x} &= \sum_{j \neq \Laughey} \inprod{V_{\Laughey}}{x}\\
  	&\leq \lambda_1 + \lambda_2 + \ldots + (-\lambda_i) + \lambda_{i+1} + \ldots + \lambda_n\\
  \end{align*}

  For $i = \Laughey$, we know that this tightness constraint is then $-1 \leq \sum_{j \neq i} \lambda_j$.
  Therefore, there must be some $\lambda_\ell \leq -1$ for the tightness constraint to be satisfied.
  However, when $i = \ell$, then we must have $V_\ell$ being able to be written as some linear combination of normals where $V_\ell = -\frac{\lambda_1}{\lambda_\ell} V_1 + \ldots + \frac{1}{\lambda_\ell} V_{\Laughey} + \ldots + (- \frac{\lambda_n}{\lambda_\ell}) V_n$.
  Since $\lambda_\ell \leq -1$, we know that $\frac{1}{\lambda_\ell} =: \lambda^*_{\Laughey} \in (0, -1]$.
  \end{proof}

  \begin{corollary}
    The abstain(1/2) property is \jessie{$\lceil \log_2 n \rceil$ embeddable, but no less?}
  \end{corollary}

  \subsection{Generalizing abstain}
  However, we know that we have to use the same set of normal for all $T_i$ for $i \in \{1,2, \ldots, n\}$.
  Therefore, if we want to write any $V_j$ as a linear combination of other normals, we can use the structure of all $T_i$ to note the possible relationships between normals.

  For example, for $x \in T_1$, we know that by the tightness constraints
  \begin{align*}
  -1 = \inprod{V_1}{x} &= \inprod{\lambda_2 V_2}{x} + \inprod{\lambda_3 V_3}{x}\\ 
  &= \lambda_2 \inprod{V_2}{x} + \lambda_3 \inprod{V_3}{x}\\
  &\geq \lambda_2 + \lambda_3
  \end{align*}

  Similarly for $x \in T_1$, we know
  \begin{align*}
  1 = \inprod{V_2}{x} &= \inprod{\lambda_1 V_1}{x} + \inprod{\lambda_3 V_3}{x}\\ 
  &= \lambda_1 \inprod{V_1}{x} + \lambda_3 \inprod{V_3}{x}\\
  &\geq -\lambda_1 + \lambda_3
  \end{align*}
  and so on for the tightness constraint on $V_3$ and $T_1$.

  Combining the inequalities, this yields
  \[
  \begin{bmatrix}
      0 & 1 & 1      \\
      -1 & 0 & 1 \\
      -1 & 1 & 0      
  \end{bmatrix}
  \begin{bmatrix}
      \lambda_1\\
      \lambda_2\\
      \lambda_3
  \end{bmatrix}
  \leq 
  \begin{bmatrix}
      -1\\
      1 \\
      1
  \end{bmatrix} 
  \]

  We can then repeat this process for $T_2$ and $T_3$ (since all three require the same set of normals) to observe

    \[
    \begin{bmatrix}
        0 & 1 & 1      \\
        -1 & 0 & 1 \\
        -1 & 1 & 0 \\ \hline
        0 & -1 & 1\\
        1 & 0 & 1\\
        1 & -1 & 0\\ \hline
        0 & 1 & -1\\
        1 & 0 &-1 \\
        1 & 1 & 0     
    \end{bmatrix}
    \begin{bmatrix}
        \lambda_1\\
        \lambda_2\\
        \lambda_3
    \end{bmatrix}
    \leq 
    \begin{bmatrix}
        -1\\
        1 \\
        1 \\ \hline
        1 \\
        -1 \\
        1\\ \hline
        1\\
        1\\
        -1\\
    \end{bmatrix} 
    \]

    We then want to find a feasible solution for this polytope.
    (It seems like $-\ones$ should be a feasible solution for all $n$, but I haven't proven it.)

    Since we know have a set of ``weights'' that show the relationship between normals, we want to consider that we then have $V_j = \sum_{\ell \neq j}\lambda_\ell V_\ell.$
    Therefore, we can set up a system (using our feasible solution $\lambda = -\ones$) by stating $V_1 = -V_2 - V_3$, $V_2 = -V_1 - V_3$, etc.

    Therefore, we have the system
    \[
    \begin{bmatrix}
        0 & -1 & -1\\
        -1 & 0 & -1 \\
        -1 & -1 & 0
    \end{bmatrix}
    \begin{bmatrix}
        V_1\\
        V_2\\
        V_3
    \end{bmatrix}
    =
    \begin{bmatrix}
        V_1\\
        V_2 \\
        V_3 
    \end{bmatrix} 
    \]
    From here, we can find (any) eigenvector corresponding to the eigenvalue $1$, and imposes some structure on the normals.

    For example, we know that $-V_1 = V_2$ by taking the Hadamard product of the eigenvector (corresponding to eigenvalue $1$) and the normals $V$.

    Once we fix $V_1$ and $V_2$, we can then look at the rest of the normals composing $V$ through the same system equality and conclude that the sum of the rest of the normals should be $\vec{0}$.
    Since $V_1 = -V_2$, their sum is also $\vec{0}$, and thus the sum of all the normals must be $\vec{0}$.

    However, a quick counterexample shows that this is not sufficient.
    Consider the matrix $V = [(0,1), (1,0), (0,-1), (-1/2, \sqrt{3} / 2), (-1/2, -\sqrt{3} / 2)]$ whose polar is given in Figure~\ref{fig:sum-to-0-gen-pos-ce}.

	\begin{figure}
	\centering
	\includegraphics[width=0.7\linewidth]{./sum-to-0-gen-pos-ce}
	\caption{Counterexample where the normals sum to $\vec{0}$ but the abstain property is not embeddable in $\reals^2$.}
	\label{fig:sum-to-0-gen-pos-ce}
	\end{figure}

	The normals are in general position and sum to $\vec{0}$, but this arrangement cannot be used to embed abstain in $\reals^2$.

	Playing around with this example, $T_i$ is not empty for all $i$, but fails to form a tight H-rep because of the adjacency requirements, which follow from the total number of $a$-faces ($a \in \{0, 1, 2, \ldots, d-1\}$) of each $T_i$.
	For example, here, two of the faces are lines, which have 4 faces in $\reals^2$, and therefore cannot make contact (i.e. have nonempty intersection with) all $T_i^j := T_i \cap \{x : \inprod{V_j}{x} = b_{i,j} \}$.
	(I'm using some implicit assumption that each $T_i^j$ is a face of $T_i$, which I believe should be true, and I think is in An Introduction to Convex Polytopes\footnote{\url{https://www.fmf.uni-lj.si/~lavric/brondsted.pdf}})

\subsection*{General geometric thoughts}
Thoughts:
\begin{itemize}
\item Can we write $B$ and $V$ so that $T = V^\circ$ and each $b_i \preceq \ones$? (componentwise inequality)?
\item There has to be some restriction on the normals stronger than them being in general position, but not as strong as them being affinely independent.
\item The matrix $V = [(0,1), (1,0), (0,-1), (-1/2, \sqrt{3} / 2), (-1/2, -\sqrt{3} / 2)]$ shows that general position and summing to $\vec{0}$ is not sufficient for the abstain property.
\item This polytope doesn't work because of something about its geometric structure.
\end{itemize}

\begin{definition}
A matrix $V$ of normals is \emph{nonredundant} \jessiet{Maybe use a different term since nonredundancy is also used in reference to properties} if \jessie{\ldots}
\end{definition}

\begin{conjecture}
The matrix $V \in \reals^{k \times d}$ forms a tight H-representation if and only if (a.) Each $T_i$ has at least $k$ faces, (b.) For all $j = \{1,2,\ldots, k\}$, the polytope $T_i^j$ is nonempty, and (c.) for all $i \in \{1,2,\ldots, n\}$ we have $|\{T_i^j\}_{j=1}^k| = k$.  That is, fixing any $T_i$, we observe $T_i^j$ is unique.
\end{conjecture}
